Supplementary material for ``Screening Sinkhorn Algorithm for regularized Optimal Transport''

### 2019-05-21 tex/v3/section_5_numeric_expes.tex
update

We have implemented our Screenkhorn in Python and used the LBFGS of
Scipy. Regarding the machine-learning based comparison, we have based our code
on the ones of POT \cite{flamary2017pot} and just replaced the Sinkhorn function call with a Screenkhorn one. We have considered the POT's default Sinkhorn stopping criterion parameters and for Screenkhorn, the LBFGS algorithm is stopped when the 
largest component of the projected gradient is smaller than $10^{-6}$, when the number of iterations of number of objective function evaluations reach $10^{5}$. For all applications, we have set $\eta=1$.

@@

We compare \emph{Screenkhorn} to Sinkhorn as implemented in Python Optimal Transport toolbox\footnote{\url{https://pot.readthedocs.io/en/stable/index.html}} on  a synthetic example. The dataset we use consists of source samples generated from a bi-dimensional gaussian mixture and target samples following the same distribution but with different gaussian means. We consider an unsupervised domain adaptation using optimal transport with entropic regularization.  Several settings are explored: different values of $\eta$, the regularization parameter, the allowed budget $\frac{n_b}{n} = \frac{m_b}{m} = $ ranging from $0.01$ to $0.99$, different values of $n$ and $m$ and, whether the cost matrix $C$ is normalized or not. We empirically measure  marginal violations $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ as well as running time expressed as $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$
Figure \ref{fig:margin_expe} summarizes the observed behaviors of both algorithms under these settings. We choose to only report results for $n=m=1000$ as we get similar findings for other values of $n$ and $m$.

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=4.55cm]{./figs/norm_M_Mu_marginals_toy_n1000} ~\hfill~
		\includegraphics[width=4.55cm]{./figs/norm_M_Mu_marginals_toy_n1000} ~\hfill \\
		\includegraphics[width=4.55cm]{./figs/norm_M_time_toy_n1000} ~\hfill~
		\includegraphics[width=4.55cm]{./figs/norm_M_div_toy_n1000}
	\end{center}
	\caption{Empirical evaluation of \emph{Screenkhorn} vs Sinkhorn for normalized cost matrix. (Top panel) marginal violations in relation with the budget of points and (Bottom panel) ratio of computation times $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$. The results are averaged over 10 trials.} 
		\label{fig:margin_expe}
\end{figure*}

\emph{Screenkhorn} provides good approximation of the marginals $\mu$ and $\nu$ for ''high'' values of the regularization parameter $\eta$ ($\eta > 1$). The approximation quality diminishes for small $\eta$. As expected $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ converge towards zero when increasing the budget of points. Remarkably those quantities are all almost null whatever the budget for high $\eta$.  According to computation gain, \emph{Screenkhorn} is, up to 2 times, faster than Sinkhorn at low budget $n_b/n$ while the reverse holds when $n_b/n$ gets close to 1.  Computational benefit of \emph{Screenkhorn} also depends on $\eta$ with appropriate values $\eta \leq 1$.

>>>>

We have implemented our Screenkhorn in Python and used the LBFGS of
Scipy. Regarding the machine-learning based comparison, we have based our code
on the ones of POT \cite{flamary2017pot} and just replaced the Sinkhorn function call with a Screenkhorn one. We have considered the POT's default Sinkhorn stopping criterion parameters and for Screenkhorn, the LBFGS algorithm is stopped when the 
largest component of the projected gradient is smaller than $10^{-6}$, when the number of iterations of number of objective function evaluations reach $10^{5}$. For all applications, we have set $\eta=1$ unless otherwise specified.

@@

We compare \emph{Screenkhorn} to Sinkhorn as implemented in Python Optimal Transport toolbox\footnote{\url{https://pot.readthedocs.io/en/stable/index.html}} on  a synthetic example. The dataset we use consists of source samples generated from a bi-dimensional gaussian mixture and target samples following the same distribution but with different gaussian means. We consider an unsupervised domain adaptation using optimal transport with entropic regularization.  Several settings are explored: different values of $\eta$, the regularization parameter, the allowed budget $\frac{n_b}{n} = \frac{m_b}{m} = $ ranging from $0.01$ to $0.99$, different values of $n$ and $m$ and, whether the cost matrix $C$ is normalized or not. We empirically measure  marginal violations as the norms $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$, running time expressed as $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$ and the relative divergence difference $\| \inr{C, P^\star} - \inr{C, P^{sc}}\|$ between \emph{Screenkhorn} and Sinkhorn.

Figure \ref{fig:margin_expe} summarizes the observed behaviors of both algorithms under these settings. We choose to only report results for $n=m=1000$ as we get similar findings for other values of $n$ and $m$.

\begin{figure*}[t]
	\begin{center}
		~\hfill~\includegraphics[width=5.5cm]{./figs/norm_M_Mu_marginals_toy_n1000}~\hfill~
		\includegraphics[width=5.5cm]{./figs/norm_M_Mu_marginals_toy_n1000}~\hfill~\\
		~\hfill~\includegraphics[width=5.5cm]{./figs/norm_M_time_toy_n1000}
		~\hfill~\includegraphics[width=5.5cm]{./figs/norm_M_div_toy_n1000}~\hfill~\\
	\end{center}
	\caption{Empirical evaluation of \emph{Screenkhorn} vs Sinkhorn for normalized cost matrix. Top panel: marginal violations in relation with the budget of points. Bottom panel: (left) ratio of computation times    $\frac{T_{\text{\emph{Screenkhorn}}}}{T_{\text{Sinkhorn}}}$ and, (right) relative divergence variation. The results are averaged over 10 trials.} 
		\label{fig:margin_expe}
\end{figure*}
%
\emph{Screenkhorn} provides good approximation of the marginals $\mu$ and $\nu$ for ''high'' values of the regularization parameter $\eta$ ($\eta > 1$). The approximation quality diminishes for small $\eta$. As expected $\norm{{\mu} -{\mu}^{\text{sc}}}_1$ and $\norm{{\nu} -{\nu}^{\text{sc}}}_1$ converge towards zero when increasing the budget of points. Remarkably marginal violations are almost negligible whatever the budget for high $\eta$.  According to computation gain, \emph{Screenkhorn} is, up to 2 times, faster than Sinkhorn at low budget $n_b/n$ while the reverse holds when $n_b/n$ gets close to 1.  Computational benefit of \emph{Screenkhorn} also depends on $\eta$ with appropriate values $\eta \leq 1$. Finally except for $\eta=0.1$ \emph{Screenkhorn} achieves a close divergence $\inr{C, P}$ compared to Sinkhorn.
